% AI Use Appendix (standalone; compile to PDF separately or append to main report)
% Compile: pdflatex ai_use_appendix
\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{array}
\usepackage{verbatim}
\usepackage[hidelinks]{hyperref}


\title{AI Use Appendix\\StockTrader Individual Assignment (Coding)}
\author{%
  Student: Ariana Rocha (afrocha)\\
  Class: 94815-A4 Agentic Technologies \\
  Assignment: HW2 Building StockTrader \\
}
\date{\today}

\begin{document}
\maketitle

\noindent\textbf{Abstract.}
This appendix describes how I used generative AI tools---\textbf{Cursor}, \textbf{ChatGPT}, and \textbf{Ollama}---while completing the StockTrader coding assignment. I steered work with written instructions and project context, reviewed every stage of output carefully, and used ChatGPT once with the rubric to stress-test the report for gaps. Ollama was required as the local LLM runtime for the assignment pipeline (strategies and evaluator), separate from my authoring workflow. Below I summarize tools, workflow, cleaned representative prompts, sample outputs, and file-level attribution (accepted, revised, or rejected).

\section{Tools I used: Cursor, ChatGPT, and Ollama}

\begin{description}
  \item[Cursor] Primary environment for repository work, code exploration, edits, and drafting. I created the GitHub repository, added a text file with the assignment description after reading it myself, and layered instructions (e.g., which agents to build, AutoGen + parallel strategies, JSON outputs, no cross-talk before evaluation). Cursor helped scaffold \texttt{src/}, \texttt{prompts/}, early \texttt{README.md} drafts, and initial report-oriented material; I reviewed and revised all of it.
  \item[ChatGPT] Used later in the process: I pasted my near-final report (or large excerpts) together with the course rubric and asked for a gap analysis---what might be missing, unclear, or weak against the rubric---not for automatic rewriting of the whole document. I incorporated only suggestions I agreed with after manual judgment.
  \item[Ollama] Part of the assignment itself: local model serving for the three \texttt{AssistantAgent} strategies plus the evaluator when running \texttt{python -m src.main}. This is runtime behavior of the system, not the same as using a chat UI to draft prose, but it is disclosed here as required AI-assisted inference on market payloads.
\end{description}

\section{Workflow: repository, steering file, and review}

\begin{enumerate}
  \item \textbf{Read the assignment} on my own, then \textbf{captured the brief} in a text file in the repo with any particulars (multi-agent StockTrader, News Sentiment Follower, Volatility Averse, Moral Trader extension, parallel execution, structured JSON, evaluator, optional backtest, Alpha Vantage news path, etc.).
  \item \textbf{Fed that context into Cursor} so subsequent edits stayed aligned with the spec; I iterated when requirements became clearer (e.g., handoff doc, Alpha Vantage hygiene, report checklist items).
  \item \textbf{Reviewed meticulously} at each stage: code paths, prompts on disk, \texttt{outputs/*.json} after runs, and narrative in report sources. I accepted, revised, or rejected suggestions explicitly rather than bulk-merging blind.
  \item \textbf{Rubric pass with ChatGPT} after I had a finished draft of the written report: attach rubric + report, ask for missing sections or weak evidence, then I edited the \texttt{.tex} / PDF pipeline myself.
\end{enumerate}

\section{Representative prompts (cleaned)}
The following are paraphrased or lightly edited from what I actually used; they are not verbatim logs.

\subsection*{Repository and build steering (Cursor / project context)}
\begin{verbatim}
I am doing HW2 StockTrader. Here is the assignment text [paste].
Build: AutoGen AssistantAgents for (1) News Sentiment Follower,
(2) Volatility Averse, (3) Moral Trader bonus. They must run in
parallel with structured JSON outputs; no strategy sees another's
output before the evaluator. Use yfinance + optional Alpha Vantage
NEWS_SENTIMENT. Include README, prompts/*.txt on disk, and outputs/
JSON per ticker plus summary.json and backtest bonus.
\end{verbatim}

\subsection*{Iteration and hygiene (Cursor)}
\begin{verbatim}
Pull from remote with rebase. Update HANDOFF so API keys only live
in .env; document Alpha Vantage rate limits per official docs.
Adjust market_data for NEWS_SENTIMENT errors and optional sort/limit.
\end{verbatim}

\subsection*{Report and documentation (Cursor)}
\begin{verbatim}
Draft README sections for setup (venv, Ollama, .env, PYTHONPATH,
run command). Match the course folder layout. Keep tone honest about
stub mode vs live Ollama for grading.
\end{verbatim}

\subsection*{Rubric gap check (ChatGPT)}
\begin{verbatim}
Here is my report draft for StockTrader HW2 [paste or attach].
Here is the rubric [paste]. Grade against the rubric as a TA would:
list gaps, missing quantitative evidence, or unclear claims only.
Do not rewrite the whole report; bullet prioritized fixes.
\end{verbatim}

\section{Representative outputs (authoring tools; not Ollama runtime)}
This subsection is about \textbf{generative assistance while building the project and the report}---what came back from \textbf{Cursor} and \textbf{ChatGPT} when I used the kinds of prompts in the previous section. It is \textbf{not} a sample of the StockTrader pipeline's on-ticker LLM outputs (Ollama strategy/evaluator text inside \texttt{outputs/*.json}); those are assignment \emph{runtime} inference and are disclosed elsewhere in this appendix. The patterns below are paraphrased or shortened; they show how the tools tended to \emph{continue} a thread---scaffolding, checks, documentation---rather than only mirroring my prompt text.

\subsection*{Repository scaffold and layout (Cursor)}
\begin{verbatim}
Proposed tree: src/main.py CLI, src/orchestration.py with asyncio.gather
for three AssistantAgent calls, prompts/*.txt, outputs/ for JSON.
Add schemas.py for Pydantic shapes; wire evaluator after strategies.
You should trace one ticker end-to-end and confirm no cross-strategy
reads before evaluation.
\end{verbatim}

\subsection*{Hygiene and secrets (Cursor; follow-up to HANDOFF / .env work)}
\begin{verbatim}
Keep ALPHAVANTAGE_API_KEY only in repo-root .env (gitignored). Optional
secrets/alphavantage_api_key.txt path is documented in secrets/README.md;
confirm .gitignore covers .env and secrets/. I can suggest a pre-commit
grep for key-like strings; you run it locally before push.
\end{verbatim}

\subsection*{README and grader-facing docs (Cursor)}
\begin{verbatim}
Draft bullets: venv activate (PowerShell), PYTHONPATH at repo root,
python -m src.main --backtest. Note STOCKTRADER_SKIP_LLM for CI vs full
Ollama run. Tone: factual, no key material in README.
\end{verbatim}

\subsection*{Rubric stress-test (ChatGPT; responds to draft + rubric paste)}
\begin{verbatim}
Prioritized gaps: (1) Tie narrative numbers to a named committed run or
outputs/*.json snapshot. (2) Say explicitly when news_features was empty.
(3) One paragraph on Moral Trader as extension vs graded pair. I am
not rewriting your PDF---address these, then tighten wording yourself.
\end{verbatim}

\noindent\textbf{How I used these outputs.} I did \textbf{not} treat any block as authoritative: I compared suggestions against the assignment, my own code reading, and CLI runs; rejected or rewrote generic or wrong steps; and separated this workflow from \textbf{Ollama}, which only runs inside \texttt{src/} for strategies and the evaluator when executing the pipeline. If a reply mostly repeated an earlier prompt, I asked for a concrete delta or stopped using that thread.

\section{File-level attribution (accept / revise / reject)}
\noindent
Disposition means how the listed artifact left my desk after review: \textbf{A}~=~accepted largely as generated (minor typos only); \textbf{R}~=~revised substantially (logic, safety, or rubric alignment); \textbf{X}~=~rejected initial suggestion and rewrote myself from spec. The rows below mix \textbf{R}, \textbf{A}, and \textbf{X} to reflect different paths---not a single default. Ollama-generated \emph{text inside} committed JSON is runtime model output, not Cursor-authored source; I still chose what to commit and when.

\begin{table}[h]
  \centering
  \footnotesize
  \setlength{\tabcolsep}{4.5pt}
  \renewcommand{\arraystretch}{1.22}
  % Three p-columns: consistent top alignment (avoids centering the Disp. column in tall rows).
  \begin{tabular}{@{}>{\raggedright\arraybackslash}p{4.35cm}>{\raggedright\arraybackslash}p{6.55cm}>{\centering\arraybackslash}p{1.45cm}@{}}
    \toprule
    Artifact & Summary (tooling touch) & Disp. \\
    \midrule
    \texttt{src/orchestration.py}, \texttt{src/main.py} & Parallel \texttt{asyncio.gather} for strategies; per-ticker flow; Cursor-assisted scaffold, manual trace of call order & R \\
    \midrule
    \texttt{src/market\_data.py} & Rejected early draft that obscured vendor failures; rewrote error handling after Alpha Vantage docs & X \\
    \midrule
    \texttt{src/evaluator.py} & Branching evaluator task text; deterministic \texttt{pattern\_note}; regex guard on contradictory ``agree/disagree'' phrasing in \texttt{analysis} & R \\
    \midrule
    \texttt{src/schemas.py}, \texttt{src/llm\_factory.py}, \texttt{src/strategies.py} & Pydantic shapes and prompt loading; thin glue modules---accepted scaffold, minor typing/import fixes only & A \\
    \midrule
    \texttt{src/backtest.py} & Replaced first Cursor heuristic with formulas aligned to assignment intent & X \\
    \midrule
    \texttt{prompts/*.txt} & Rejected generic first drafts; rewrote all four for structured outputs and rubric tone & X \\
    \midrule
    \texttt{README.md}, \texttt{src/README.md} & Cursor setup draft; accepted after command-by-command verification and tone trim & A \\
    \midrule
    \texttt{docs/*.md} (e.g.\ \texttt{ALPHA\_VANTAGE.md}, \texttt{HANDOFF\_NEXT\_AGENT.md}) & Operational notes; fact-checked against primary sources; light copy edits & R \\
    \midrule
    \texttt{report/comparative\_analysis.tex} & Reorganized report outline and TikZ against committed \texttt{outputs/}; not the first AI-suggested structure & X \\
    \midrule
    \texttt{report/ai\_use\_appendix.tex} & Disclosure prose authored by hand; LaTeX table formatting iterated locally & R \\
    \midrule
    \texttt{report/report\_checklist.md} & Personal checklist; occasional wording nits only & R \\
    \midrule
    \texttt{report/LATEX\_BUILD.md} & One-page build commands; accepted as-is after path check & A \\
    \midrule
    \texttt{report/walkthrough/*.html}, \texttt{report/walkthrough/walkthrough.css} & Map, canvases, market-data page; iterated layout/CSS for grader clarity & R \\
    \midrule
    \texttt{report/walkthrough/Walkthrough\_Artifacts.html} & Outputs walkthrough page; split from map for readability; refined SVG + file cards locally & R \\
    \midrule
    \texttt{scripts/*.py} & Redaction / maintenance scripts; small, reviewed line-by-line; kept nearly verbatim & A \\
    \midrule
    \texttt{outputs/*.json} & Produced by \texttt{python -m src.main} (Ollama or stub); I diffed keys, spot-checked decisions vs.\ payload, chose commit set for grading & Runtime \\
    \bottomrule
  \end{tabular}
\end{table}

\section{What I verified independently}
I treated anything that touches grades or keys as my responsibility, not the model's. Independently (without relying on tool summaries) I confirmed the following by reading code, running the CLI, and opening files:

\begin{itemize}
  \item \textbf{Schema and JSON shape.} Each per-ticker file includes \texttt{strategy\_a|b|c} with \texttt{decision}, \texttt{confidence}, \texttt{justification}, plus an \texttt{evaluator} object with \texttt{agents\_agree}, \texttt{analysis}, and \texttt{pattern\_note}, matching \texttt{src/schemas.py} and the assignment's structured-output requirement.
  \item \textbf{Parallelism and isolation.} In \texttt{src/orchestration.py}, the three strategies are awaited together (\texttt{asyncio.gather}) on the same market payload; nothing in that phase passes one strategy's completion text into another's system message. The evaluator is only called after all three structured results exist for that ticker.
  \item \textbf{Deterministic routing metadata.} \texttt{pattern\_note} is derived in Python from the three decisions (\texttt{\_pattern\_note\_from\_decisions} in \texttt{src/evaluator.py}), so the saved note cannot drift from the actual votes even if the LLM paraphrases poorly.
  \item \textbf{Evaluator vs.\ branch consistency.} When all three decisions match, \texttt{agents\_agree} is \texttt{true}; when the set of labels has more than one value, it is \texttt{false}. I stepped through branches in \texttt{evaluator\_task} and spot-checked tickers where News and Vol disagree.
  \item \textbf{Contradiction guard.} I read the regex blocks in \texttt{run\_evaluator} that replace evaluator prose when the model claims disagreement under unanimity (or the reverse), and confirmed they fire only on the intended edge cases.
  \item \textbf{Market payload behavior.} \texttt{build\_market\_payload} and Alpha Vantage paths: I verified empty \texttt{articles}, null sentiment, and vendor error/rate-limit text surface under \texttt{news\_features} without crashing the pipeline, and that strategies still receive a single coherent JSON user message.
  \item \textbf{Stub vs.\ live path.} With \texttt{STOCKTRADER\_SKIP\_LLM=1}, the pipeline returns stub structured objects so the graph can be tested without Ollama; I confirmed that path in \texttt{evaluator.py} and strategy code paths as documented in README.
  \item \textbf{Report numbers vs.\ artifacts.} Tabulated or narrative statistics in \texttt{comparative\_analysis.tex} were cross-checked against committed \texttt{outputs/*.json} and run dates so I did not accidentally describe an old run.
  \item \textbf{Secrets and repo hygiene.} Confirmed API keys are not committed; \texttt{.env.example} documents variables; assignment handoff docs describe rate limits from primary vendor documentation.
\end{itemize}

\section{Weak spots, failure modes, and mitigations}
Generative components remain the main risk surface. These are the issues I actually saw or explicitly guarded against; they are not hypothetical:

\begin{itemize}
  \item \textbf{Sparse news + strong price/vol fields.} Small models sometimes overweight narrative phrasing (``turbulence'', ``risk'') relative to numeric gates in the JSON, or invent nuance not tied to a field. \textbf{Mitigation:} prompt text requires citations to concrete keys; Volatility prompt asks for plain-English lead sentences tied to thresholds; evaluator branch text forces substantive comparison on disagreement.
  \item \textbf{Evaluator prose contradicting arithmetic unanimity.} The model occasionally emitted words like ``disagree'' when all three \texttt{decision} strings matched (or claimed full agreement when they did not). \textbf{Mitigation:} post-process replacement strings in \texttt{run\_evaluator} after structured extraction; \texttt{pattern\_note} overwritten from code, not the LLM.
  \item \textbf{Extension strategy (Moral) variability.} Third philosophy is useful but less constrained by external data; confidence and justification can look mismatched (e.g.\ low confidence with assertive wording). \textbf{Mitigation:} kept scope clearly ``same JSON'' and relied on evaluator to contrast philosophies without treating Moral as ground truth.
  \item \textbf{Vendor fragility.} Alpha Vantage rate limits and error payloads appear inside \texttt{news\_features.note}; strategies must not treat that blob as tradable news. \textbf{Mitigation:} documented in \newline \texttt{docs/ALPHA\_VANTAGE.md}; optional redaction script under \texttt{scripts/} for committed samples.
  \item \textbf{Committed JSON drift.} Any snapshot in \texttt{outputs/} can become stale relative to code after refactors. \textbf{Mitigation:} re-ran pipeline when logic changed; verified evaluator and strategy keys before final submission.
\end{itemize}

\end{document}
